Identifying Heterogeneity in SAR Data with New Test Statistics

### 2024-04-11 Text/Identifying Heterogeneity in SAR Data with New Test Statistics.Rmd
Changes added in Rmd file

# author citation list in chicago format
authorcitation: |
  Frery, A.C.; Alpala, J.; Nascimento,A.D.C.
# firstnote to eighthnote
#firstnote: |
 # Current address: Updated affiliation
#secondnote: |
  #These authors contributed equally to this work.
correspondence: |
  alejandro.frery@vuw.ac.nz
# document options
journal: remotesensing
type: article
status: submit
# front matter
#simplesummary: |
  #A Simple summary goes here.
abstract: |
  This paper presents a statistical approach to identify the underlying roughness characteristics in synthetic aperture radar (SAR) intensity data. The physical modeling of this kind of data allows to use the Gamma distribution in the presence of fully-developed speckle, i.e., when there are infinitely many independent backscatterers per resolution cell. This situation is often called "homogeneous" or 
  "textureless" areas. The $\mathcal{G}_I^0$ distribution is also a widely accepted law for heterogeneous and extremely heterogeneous areas. We propose three test statistics for discriminating between homogeneous and non-homogeneous areas, i.e., between Gamma and GI0 distributed data, both with a known number of looks. The first test statistic utilizes a bootstrapped non-parametric estimator of the Shannon entropy, offering a robust evaluation in the face of uncertain distributional assumptions. The second test uses the classic coefficient of variation. The third test employs an alternative form of estimating the CV based on the ratio of the mean absolute deviation from the median to the median. We apply our test statistic to produce maps of $p$-values for the hypothesis of homogeneity. These maps identify different types of targets. We show that our proposal outperforms its competitors using simulated and SAR data.
# back matter
keywords: |
  SAR; heterogeneity; entropy;  coefficient of variation; hypothesis tests
acknowledgement: |
  This study was financed in part by the Coordenação de Aperfeiçoamento de Pessoal de Nível Superior - Brasil (CAPES) - Finance Code 001.
authorcontributions: |
  For research articles with several authors, a short paragraph specifying their 
  individual contributions must be provided. The following statements should be 
  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
  limited to those who have contributed substantially to the work reported.
funding: |
  Please add: "This research received no external funding" or "This research 
  was funded by NAME OF FUNDER grant number XXX."  and ``The APC was funded 
  by XXX''. Check carefully that the details given are accurate and use the 
  standard spelling of funding agency names at 
  \url{https://search.crossref.org/funding}, any errors may affect your future 
  funding.
institutionalreview: |
 Not applicable.
#informedconsent: |
  # Any research article describing a study involving humans should contain this 
  # statement. Please add ``Informed consent was obtained from all subjects 
  # involved in the study.'' OR ``Patient consent was waived due to REASON 
  # (please provide a detailed justification).'' OR ``Not applicable'' for 
  # studies not involving humans. You might also choose to exclude this statement 
  # if the study did not involve humans.

  # Written informed consent for publication must be obtained from participating 
  # patients who can be identified (including by the patients themselves). Please 
  # state ``Written informed consent has been obtained from the patient(s) to 
  # publish this paper'' if applicable.
dataavailability: |
  Not applicable.
conflictsofinterest: |
  The authors declare no conflict of interest.
supplementary: |
 This article is fully reproducible using RMarkdown. All code and data are accessible at
 \url{https://github.com/rjaneth/identifying-heterogeneity-in-sar-data-with-new-test-statistics} (accessed on 5 April 2024).
# abbreviations:
#   - short: MDPI
#     long: Multidisciplinary Digital Publishing Institute
#   - short: DOAJ
#     long: Directory of open access journals
#   - short: TLA
#     long: Three letter acronym
#   - short: LD 
#     long: linear dichroism

header-includes:
   - \usepackage[english]{babel}
   - \usepackage{bm,bbm}
   - \usepackage{mathrsfs}
   - \usepackage{siunitx}
   - \usepackage{graphicx}
   - \usepackage{url}
   - \usepackage[T1]{fontenc}
   - \usepackage{polski}
   - \usepackage{booktabs}
   - \usepackage{color}
   - \usepackage{amsmath}
   - \usepackage{multirow}
   - \usepackage{subcaption}
   - \captionsetup[sub]{labelfont=bf,textfont=bf}

bibliography: references.bib
#appendix: appendix.tex
endnotes: false
output: 
  rticles::mdpi_article:
    extra_dependencies: longtable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE)
#knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(reshape2)
#library(plotly)
library(knitr)
library(pandoc)
library(gridExtra)
library(gtools)
library(stats4)
library(rmutil)
library(scales)
library(tidyr)
library(rmutil)
library(invgamma)
library(tidyverse)
library(RColorBrewer)
library(ggsci)
library(ggpubr)
library(patchwork)
library(dplyr)
#options(kableExtra.latex.load_packages = FALSE)
#library(devtools)
#devtools::install_github("haozhu233/kableExtra")
library(kableExtra)
library(ggthemes)
library(latex2exp)
library(e1071)# kurtosis
library(viridis)
library(nortest)# AD
library(MASS)#
library(fitdistrplus)# model selection AIC and BIC

@@

# if(!require("rstudioapi")) install("rstudioapi")
# setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# opts_chunk$set(comment="",
#                message=FALSE,
#                tidy.opts=list(keep.blank.line=TRUE, width.cutoff=120),
#                options(width=100), 
#                # cache=TRUE,
#                fig.align='center',
#                fig.height=6, fig.width=10,
#                fig.show='hold',
#                cache=TRUE)

@@

```{r DefinitionSimulationRoutines, echo=FALSE}
#set.seed(1234567890, kind="Mersenne-Twister")
```
\setlength{\tabcolsep}{6pt} 
\newcommand{\bias}{\operatorname{Bias}}
\newcommand{\widebar}[1]{\overline{#1}}

@@

Synthetic Aperture Radar (SAR) technology has become an important tool for environmental monitoring and disaster management, providing valuable imagery in different conditions, including day or night and diverse weather situations&nbsp;\cite{Moreira2013,Mu2019}.
However, the effective utilization of SAR data depends on a thorough understanding of its statistical properties, especially the impact of speckle effects, which are inherent in SAR data due to the coherent nature of the imaging process&nbsp;\cite{Argenti2013}.

In the presence of non-Gaussian speckle interference, SAR data requires a reliable statistical model for accurate processing. The $\mathcal{G}^0$ distribution, suitable for SAR data, includes the Gamma law as a limit case for fully developed speckle&nbsp;\cite{Ferreira2020}, providing flexibility with fewer parameters for analysis.

Our work is motivated to improve the identification of potential roughness features in SAR intensity data. 
Physical modeling of SAR data allows to take advantage of the gamma distribution in the presence of fully developed speckle, creating a scenario with an infinite number of independent backscatterers per unit of resolution, often referred to as homogeneous areas.

In this context, we present a set of three novel test statistics designed to discriminate between homogeneous and non-homogeneous areas, specifically aimed at differentiating between Gamma and $\mathcal{G}^0$ distributed data, each with a known number of appearances. 
Hence, to understand the core ideas behind our approach, we explore into the use of properties such as entropy and coefficient of variation.

The entropy is a fundamental concept in information theory with broad applications to pattern recognition, statistical physics, stochastic dynamics, and statistics. 
Shannon introduced it for a random variable in 1948&nbsp;\cite{Shannon1948} as a measure of information and uncertainty.  
In statistics, Shannon entropy is a crucial descriptive parameter, particularly for assessing data dispersion and conducting tests for normality, exponentiality, and uniformity&nbsp;\cite{Wieczorkowski1999,Zamanzade2012}. 
Entropy estimation involves several practical difficulties, especially when the model is unknown; in these cases, non-parametric methods are used.

Spacing methods have been discussed as a non-parametric approach in&nbsp;\cite{AlizadehNoughabi2010,Subhash2021}. 
This strategy offers flexibility to address a wide range of models without imposing specific parametric restrictions.

In our study, entropy serves as the fundamental parameter, providing a robust assessment by means of a bootstrapped non-parametric estimator of Shannon entropy.

Simultaneously, we use the classic coefficient of variation (CV) and an alternative form based on the ratio between the median absolute deviation and the median. 
The CV is known as an efficient and robust index of textural information, measures the homogeneity of the image&nbsp;\cite{Lopes1990}.

The CV, introduced by Pearson&nbsp;\cite{Pearson1896} in 1896, is a relative measure of dispersion widely used in various areas of applied statistics including sampling, biostatistics, medical and biological research, climatology, and other fields \cite{hendricks1936sampling, Tian2005,SubrahmanyaNairy2003,Chankham2024}.
It facilitates the comparison of variability between different populations and is especially valuable for comparing variables with different units. 
This is because when the main purpose is to compare variations of several variables, the standard deviation cannot serve as an adequate measure of variation unless all variables are expressed in the same unit of measurement and have identical mean values. 
When these requirements are not hold for the variables under study, CV is the relative measure typically used in real applications. 
The variable with the highest CV value is the one with the highest relative dispersion around the mean&nbsp;\cite{Banik2011}.

@@

The article is structured as follows: 
Section&nbsp;\ref{sec:Background} covers statistical modeling and entropy estimation for Intensity SAR data. Section&nbsp;\ref{sec:test} outlines hypothesis testing based on non-parametric entropy estimators and and coefficient of variation. 
In Section&nbsp;\ref{sec:Results}, we present experimental results. 
Finally, in Section&nbsp;\ref{sec:conclusion} conclusions are exhibited.

@@

The primary models used for intensity SAR data include the Gamma and $\mathcal{G}_I^0$  distributions&nbsp;[@Frery1997]. 
The first is suitable for fully developed speckle and is a limiting case of the second, which is appealing due to its versatility in accurately representing regions with various roughness characteristics&nbsp;[@Cassetti2022].
We denote $Z \sim \Gamma_{\text{SAR}}(L, \mu)$ and $Z \sim G_I^0(\alpha, \gamma, L)$ to indicate that $Z$ follows the distributions characterized by the respective probability density functions:
\begin{align}
	f_Z(z;L, \mu)&=\frac{L^L}{\Gamma(L)\mu^L}z^{L-1}\exp\left\{-Lz/\mu\right\} \mathbbm 1_{\mathbbm R_+}(z),\label{E:gamma1}\\
	f_Z(z; \alpha, \gamma, L)&=\frac{L^L\Gamma(L-\alpha)}{\gamma^{\alpha}\Gamma(-\alpha)\Gamma(L)}\cdot\frac{z^{L-1}}{(\gamma+Lz)^{L-\alpha}} \mathbbm 1_{\mathbbm R_+}(z),\label{E:gi01}
\end{align}
where, in&nbsp;\eqref{E:gamma1} $\mu > 0$ is the mean; in&nbsp;\eqref{E:gi01}  $\gamma > 0$ is the scale, $\alpha < -1$ measures the roughness,  $L \geq 1$ is the number of looks, $\Gamma(\cdot)$ is the gamma function, and $\mathbbm 1_{A}(z)$ is the indicator function of the set $A$.

From \eqref{E:gi01}, the $r$th moment of $Z$ is expressed as:
\begin{align}
	E_{G_I^0}\left(Z^r\right)=\left(\frac{\gamma}{L}\right)^r\frac{\Gamma(-\alpha-r)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+r)}{\gamma(L)}, \quad \alpha <-r. 
	\label{E:rmom}
\end{align}

Even though the $\mathcal{G}_I^0$  distribution is defined by the parameters $\alpha$ and $\gamma$, SAR literature commonly utilizes the texture $\alpha$ and the mean $\mu$&nbsp;\cite{Nascimento2010}.
In this way, we compute the expected value $\mu$ using the expression in&nbsp;\eqref{E:rmom}, and we reparametrize&nbsp;\eqref{E:gi01} using $\mu$, $\alpha$, and $L$. Then
\begin{align*}
	\mu=\left(\frac{\gamma}{L}\right)\frac{\Gamma(-\alpha-1)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+1)}{\gamma(L)}=-\frac{\gamma}{\alpha+1}.
\end{align*}
Thus, the probability density function that characterize the $G_I^0(\mu, \alpha, L)$ law is
\begin{equation}
		f_Z(z; \mu, \alpha, L)=\frac{L^L\Gamma(L-\alpha)}{\big(-\mu(\alpha+1)\big)^{\alpha}\Gamma(-\alpha)\Gamma(L)} \frac{z^{L-1}}{\big(-\mu(\alpha+1)+Lz\big)^{L-\alpha}}.\label{E:gi02}
\end{equation}

## The Shannon Entropy
The parametric representation of Shannon entropy for a system described by a continuous random variable is:
\begin{equation}
  \label{E:entropy2}
  H(Z)=-\int_{-\infty }^\infty \ f(z)\ln f(z)\, \mathrm{d}z,
\end{equation}
here, $f(\cdot)$ is the probability density function that characterizes the distribution of the real-valued random variable $Z$.

Using&nbsp;\eqref{E:entropy2}, we can express the Shannon entropy of $\Gamma_{\text{SAR}}$in&nbsp;\eqref{E:gamma1} and $G_I^0$ in&nbsp;\eqref{E:gi02} based on&nbsp;\cite{Cassetti2022} and&nbsp;\cite{Ferreira2020}:
\begin{equation}
\label{E:E-gamma}
H_{\Gamma_{\text{SAR}}}(L, \mu) =   L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) + \ln \mu, 
\end{equation}
\begin{multline}
\label{E:E-GIO}
H_{G_I^0}(\mu, \alpha, L) =L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) +\ln \mu -\ln\Gamma(L-\alpha)\\
+ (L-\alpha) \psi^{(0)}(L-\alpha)-(1-\alpha)\psi^{(0)}(-\alpha)+\ln (-1-\alpha)+\ln\Gamma(-\alpha)-L,
\end{multline}
where $\psi^{(0)}(\cdot)$ is the digamma function. Figure&nbsp;\ref{fig:Plot_GI0_to_gamma1}, shows the behaviour of the entropy of \(G_I^0\) when when \(\alpha \in \left\{-\infty, -20, -8, -3\right\}\), observing its convergence towards the entropy of \(\Gamma_{\text{SAR}}\) as \(\alpha\) takes large negative values.

@@

The problem of non-parametric estimating of $H(Z)$ has been considered by many authors including&nbsp;\cite{vasicek1976test, Wieczorkowski1999,correa1995new,AlOmari2019}, who proposed estimators based on spacings.

One of the earliest non-parametric estimators relying on spacings was introduced by Vasicek&nbsp;\cite{vasicek1976test}. 
Assuming that $\bm{Z}=(Z_1, Z_2,\ldots,Z_n)$ is a random sample from the distribution $F(z)$, the estimator is defined as:
\begin{equation*}
\label{E:Vas}
	\widehat{H}_{\text{V}}(\bm{Z})=\frac{1}{n}\sum_{i=1}^{n}\ln\left[\frac{n}{2m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
	\end{equation*}
where $m<n/2$ is a positive integer, $Z_{(i+m)}-Z_{(i-m)}$ is the $m$-spacing and $Z_{(1)}\leq Z_{(2)}\leq\ldots\leq Z_{(n)}$ are the order statistics and $Z_{(i)}= Z_{(1)}$ if $i<1$, $Z_{(i)}= Z_{(n)}$ if $i>n$.

Several authors have explored adaptations to Vasicek's estimator. 
We consider three estimators known for their superior performance&nbsp;\cite{Cassetti2022}:

* Correa&nbsp;\cite{correa1995new}: $\widehat{H}_{\text{C}}$.

* Ebrahimi et al&nbsp;\cite{Ebrahimi1994}: $\widehat{H}_{\mathrm{E}}$.

* Al Omary&nbsp;\cite{IbrahimAlOmari2014}: $\widehat{H}_{\mathrm{AO}}$.

@@

We employ the bootstrap technique to refine the precision of existing non-parametric entropy estimators. 
This approach involves generating new datasets through resampling with repetition from an existing one&nbsp;\cite{Michelucci2021}.

<!-- This approach involves generating new datasets through resampling with repetition from an existing one. -->

>>>>

# author citation list in chicago format
authorcitation: |
  Frery, A.C.; Alpala, J.; Nascimento,A.D.C.
# firstnote to eighthnote
#firstnote: |
 # Current address: Updated affiliation
#secondnote: |
  #These authors contributed equally to this work.
correspondence: |
  alejandro.frery@vuw.ac.nz
# document options
journal: remotesensing
type: article
status: submit
# front matter
#simplesummary: |
  #A Simple summary goes here.
abstract: |
  This paper presents a statistical approach to identify the underlying roughness characteristics in synthetic aperture radar (SAR) intensity data. 
  The physical modeling of this kind of data allows to use the Gamma distribution in the presence of fully-developed speckle, i.e., when there are infinitely many independent backscatterers per resolution cell. This situation is often called "homogeneous" or  "textureless" regions. 
  The $\mathcal{G}_I^0$ distribution is also a widely accepted law for heterogeneous and extremely heterogeneous regions. We propose three test statistics to distinguish between homogeneous and inhomogeneous regions, i.e. between gamma and \(\mathcal{G}_I^0\) distributed data, both with a known number of looks. 
  The first test statistic uses a bootstrapped nonparametric estimator of Shannon entropy, which provides a robust assessment in the face of uncertain distributional assumptions.
  The second test uses the classical coefficient of variation. 
  The third test uses an alternative form of estimating the CV based on the ratio of the mean absolute deviation from the median to the median. 
  We apply our test statistic to create maps of \(p\) values for the hypothesis of homogeneity. These maps identify different types of targets. 
  We show that our proposal outperforms its competitors using simulated and SAR data.
keywords: |
  SAR; heterogeneity; entropy;  coefficient of variation; hypothesis tests
acknowledgement: |
  This study was financed in part by the Coordenação de Aperfeiçoamento de Pessoal de Nível Superior - Brasil (CAPES) - Finance Code 001.
authorcontributions: |
  For research articles with several authors, a short paragraph specifying their 
  individual contributions must be provided. The following statements should be 
  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
  limited to those who have contributed substantially to the work reported.
funding: |
  Please add: "This research received no external funding" or "This research 
  was funded by NAME OF FUNDER grant number XXX."  and ``The APC was funded 
  by XXX''. Check carefully that the details given are accurate and use the 
  standard spelling of funding agency names at 
  \url{https://search.crossref.org/funding}, any errors may affect your future 
  funding.
institutionalreview: |
 Not applicable.
#informedconsent: |
  # Any research article describing a study involving humans should contain this 
  # statement. Please add ``Informed consent was obtained from all subjects 
  # involved in the study.'' OR ``Patient consent was waived due to REASON 
  # (please provide a detailed justification).'' OR ``Not applicable'' for 
  # studies not involving humans. You might also choose to exclude this statement 
  # if the study did not involve humans.

  # Written informed consent for publication must be obtained from participating 
  # patients who can be identified (including by the patients themselves). Please 
  # state ``Written informed consent has been obtained from the patient(s) to 
  # publish this paper'' if applicable.
dataavailability: |
  Not applicable.
conflictsofinterest: |
  The authors declare no conflict of interest.
supplementary: |
 This article is fully reproducible using RMarkdown. All code and data are accessible at
 \url{https://github.com/rjaneth/identifying-heterogeneity-in-sar-data-with-new-test-statistics} (accessed on 12 April 2024).
# abbreviations:
#   - short: MDPI
#     long: Multidisciplinary Digital Publishing Institute
#   - short: DOAJ
#     long: Directory of open access journals
#   - short: TLA
#     long: Three letter acronym
#   - short: LD 
#     long: linear dichroism

header-includes:
   - \usepackage[english]{babel}
   - \usepackage{bm,bbm}
   - \usepackage{mathrsfs}
   - \usepackage{siunitx}
   - \usepackage{graphicx}
   - \usepackage{url}
   - \usepackage[T1]{fontenc}
   - \usepackage{polski}
   - \usepackage{booktabs}
   - \usepackage{color}
   - \usepackage{xcolor}
   - \usepackage{amsmath}
   - \usepackage{multirow}
   - \usepackage{subcaption}
   - \captionsetup[sub]{labelfont=bf,textfont=bf}

bibliography: references.bib
#appendix: appendix.tex
#endnotes: false
output:
  rticles::mdpi_article:
    extra_dependencies: longtable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE)
#knitr::opts_chunk$set(echo = FALSE)
library(fitdistrplus)
library(ggplot2)
library(reshape2)
#library(plotly)
library(knitr)
library(pandoc)
library(gridExtra)
library(gtools)
library(stats4)
library(rmutil)
library(scales)
library(tidyr)
library(rmutil)
library(invgamma)
library(tidyverse)
library(RColorBrewer)
library(ggsci)
library(ggpubr)
library(patchwork)
library(dplyr)
#options(kableExtra.latex.load_packages = FALSE)
#library(devtools)
#devtools::install_github("haozhu233/kableExtra")
library(kableExtra)
library(ggthemes)
library(latex2exp)
library(e1071)# kurtosis
library(viridis)
library(nortest)# AD
#library(MASS)#
# model selection AIC and BIC

@@

@@

\newcommand{\bias}{\operatorname{Bias}}
\newcommand{\widebar}[1]{\overline{#1}}

@@

Synthetic Aperture Radar (SAR) technology has become an important tool for environmental monitoring and disaster management. It provides valuable images under various conditions, including day or night and different weather situations&nbsp;\cite{Moreira2013,Mu2019}.
However, the effective use of SAR data depends on a thorough understanding of its statistical properties, especially the impact of speckle effects inherent in SAR data due to the coherent nature of the imaging process&nbsp;\cite{Argenti2013}.

In the presence of non-Gaussian speckle interference, SAR data requires a reliable statistical model for accurate processing. 
The $\mathcal{G}^0$ distribution, which is suitable for SAR data, includes the
Gamma law as the limiting case for fully developed speckle&nbsp;\cite{Ferreira2020} and provides flexibility with fewer parameters for analysis.

Our work aims to improve the identification of potential roughness features in SAR intensity data. 
Physical modeling of SAR data allows to take advantage of the gamma distribution in the presence of fully developed speckles and to create a scenario with an infinite number of independent backscatterers per resolution unit, often referred to as homogeneous regions.

In this context, we present a set of three novel test statistics that aim to distinguish between homogeneous and non-homogeneous domains, in particular between gamma and \(\mathcal{G}^0\) distributed data, each with a known number of looks. 
To understand the core ideas behind our approach, we explore the use of properties such as entropy and coefficient of variation.

Entropy is a fundamental concept in information theory with far-reaching applications in pattern recognition, statistical physics, image processing, edge detection and SAR image analysis&nbsp;\cite{Presse2013,MohammadDjafari2015,Avval2021, Nascimento2014,Nascimento2019}. 
Shannon introduced it in 1948&nbsp;\cite{Shannon1948} for a random variable as a measure of information and uncertainty.
In statistics, Shannon entropy is a crucial descriptive parameter, especially for evaluating data dispersion and performing tests for normality, exponentiality and uniformity&nbsp;\cite{Wieczorkowski1999,Zamanzade2012}. 
Entropy estimation is associated with some practical difficulties, especially when the model is unknown. In these cases, non-parametric methods are used.
Spacing methods have been discussed as a non-parametric approach in&nbsp;\cite{AlizadehNoughabi2010,Subhash2021}. This strategy provides the flexibility to consider a wide range of models without enforcing certain parametric constraints.

The coefficient of variation (CV), introduced in 1896 by Pearson&nbsp;\cite{Pearson1896}, is a relative dispersion measure widely used in various fields of applied statistics, including sampling, biostatistics, medical and biological research, climatology and other fields&nbsp;\cite{hendricks1936sampling,Tian2005,SubrahmanyaNairy2003,Chankham2024}.
It facilitates the comparison of variability between different populations and is particularly valuable for comparing variables with different units. 
This is because when the main purpose is to compare the variations of several variables, the standard deviation can only serve as an adequate measure of variation if all variables are expressed in the same unit of measurement and have identical means. 
If these conditions are not met for the variables under investigation, the CV is the relative measure that is usually used in real applications. The variable with the highest CV value is the one with the largest relative dispersion around the mean value&nbsp;\cite{Banik2011}.

In our study, entropy serves as a fundamental parameter that allows a robust evaluation using a bootstrapped nonparametric estimator of the Shannon entropy.
At the same time, we use the classical CV and an alternative form based on the ratio between the median absolute deviation and the median. 
The CV is known as an efficient and robust index for texture information and measures the homogeneity of the image&nbsp;\cite{Lopes1990}.
Applying these test statistics to generate maps of \(p\) values facilitates testing the homogeneity hypothesis and reveals different types of targets in the SAR data. 
Through extensive experiments with real and simulated SAR data, we demonstrate the superiority of our proposed method over existing approaches.

@@

The article is structured as follows: Section&nbsp;\ref{sec:Background}
deals with statistical modeling and entropy estimation for intensity SAR
data. 
Section&nbsp;\ref{sec:test} outlines hypothesis tests based on nonparametric entropy estimators and coefficients of variation. 
In Section&nbsp;\ref{sec:Results} we present experimental results. 
Finally, in Section&nbsp;\ref{sec:conclusion} the conclusions are presented.

@@

The primary models used for intensity SAR data include the Gamma and $\mathcal{G}_I^0$  distributions&nbsp;[@Frery1997]. 
The first is suitable for fully developed speckle and is a limiting case of the second model, which is interesting due to its versatility in accurately representing regions with different roughness properties&nbsp;[@Cassetti2022].
We denote $Z \sim \Gamma_{\text{SAR}}(L, \mu)$ and $Z \sim \mathcal{G}_I^0(\alpha, \gamma, L)$ to indicate that $Z$ follows the distributions characterized by the respective probability density
functions (pdfs): 
\begin{align}
    f_Z(z;L, \mu)&=\frac{L^L}{\Gamma(L)\mu^L}z^{L-1}\exp\left\{-Lz/\mu\right\} \mathbbm 1_{\mathbbm R_+}(z)\label{E:gamma1}\\
    \intertext{ and }
    f_Z(z; \alpha, \gamma, L)&=\frac{L^L\Gamma(L-\alpha)}{\gamma^{\alpha}\Gamma(-\alpha)\Gamma(L)}\cdot\frac{z^{L-1}}{(\gamma+Lz)^{L-\alpha}} \mathbbm 1_{\mathbbm R_+}(z),\label{E:gi01}
\end{align} 
where \(\mu > 0\) is the mean, \(\gamma > 0\) is the scale, \(\alpha < -1\) measures
the roughness, \(L \geq 1\) is the number of looks, \(\Gamma(\cdot)\) is
the gamma function, and \(\mathbbm 1_{A}(z)\) is the indicator function
of the set \(A\).

If we take the law with the density&nbsp;\eqref{E:gi01}, the \(r\)th non-central moment of \(Z\) is expressed as follows:
\begin{align}
    E_{G_I^0}\left(Z^r\right)=\left(\frac{\gamma}{L}\right)^r\frac{\Gamma(-\alpha-r)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+r)}{\Gamma(L)}, \quad \alpha <-r. 
    \label{E:rmom}
\end{align}

Although the \(\mathcal{G}_I^0\) distribution is defined by the parameters \(\alpha\) and \(\gamma\), in the SAR literature&nbsp;\cite{Nascimento2010} the texture \(\alpha\) and the mean \(\mu\) are usually used. In this way, we calculate the expected value \(\mu\) using the expression in&nbsp;\eqref{E:rmom}, and we reparametrize&nbsp;\eqref{E:gi01} with \(\mu\),
\(\alpha\), and \(L\).
Then \begin{align*}
    \mu=\left(\frac{\gamma}{L}\right)\frac{\Gamma(-\alpha-1)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+1)}{\gamma(L)}=-\frac{\gamma}{\alpha+1}.
\end{align*} 
The reparametrized pdf that characterizes the law \(G_I^0(\mu, \alpha, L)\) is thus
\begin{equation}
        f_Z(z; \mu, \alpha, L)=\frac{L^L\Gamma(L-\alpha)}{\big[-\mu(\alpha+1)\big]^{\alpha}\Gamma(-\alpha)\Gamma(L)} \frac{z^{L-1}}{\big[-\mu(\alpha+1)+Lz\big]^{L-\alpha}}.\label{E:gi02}
\end{equation}

## The Shannon Entropy
The parametric representation of Shannon entropy for a system described by a continuous random variable is:
\begin{equation}
  \label{E:entropy2}
  H(Z)=-\int_{-\infty }^\infty \ f(z)\ln f(z)\, \mathrm{d}z,
\end{equation}
here, \(f(\cdot)\) is the pdf that characterizes the distribution of the real-valued random variable \(Z\).

Using&nbsp;\eqref{E:entropy2}, we can express the Shannon entropy of $\Gamma_{\text{SAR}}$in&nbsp;\eqref{E:gamma1} and $G_I^0$ in&nbsp;\eqref{E:gi02} based on&nbsp;\cite{Cassetti2022} and&nbsp;\cite{Ferreira2020}:
\begin{equation}
\label{E:E-gamma}
H_{\Gamma_{\text{SAR}}}(L, \mu) =   L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) + \ln \mu, 
\end{equation}
\begin{multline}
\label{E:E-GIO}
H_{G_I^0}(\mu, \alpha, L) =L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) +\ln \mu -\ln\Gamma(L-\alpha)\\
+ (L-\alpha) \psi^{(0)}(L-\alpha)-(1-\alpha)\psi^{(0)}(-\alpha)+\ln (-1-\alpha)+\ln\Gamma(-\alpha)-L,
\end{multline}
where $\psi^{(0)}(\cdot)$ is the digamma function. Figure&nbsp;\ref{fig:Plot_GI0_to_gamma1}, shows the behavior of the entropy of \(G_I^0\) when \(\alpha \in \left\{-\infty, -20, -8, -3\right\}\), where the approximation to the entropy of \(\Gamma_{\text{SAR}}\) can be observed when \(\alpha\) assumes large negative values.
In an interpretative sense, we can conclude that the more heterogeneous (high values of $\alpha$) the SAR region is, the higher the value of entropy (or degree of disorder).
It is worth noting that this conclusion confirms what is written in the SAR literature.

@@

The problem of non-parametric estimating of $H(Z)$ has been considered by many authors including&nbsp;\cite{vasicek1976test, Wieczorkowski1999,correa1995new,AlOmari2019}, who have proposed estimators based on spacings.

One of the first non-parametric estimators based on spacings was introduced by Vasicek&nbsp;\cite{vasicek1976test}. 
Under the assumption that \(\bm{Z}=(Z_1, Z_2,\ldots,Z_n)\) is a random sample from the distribution \(F(z)\), the estimator is defined as:
\begin{equation*}
\label{E:Vas}
	\widehat{H}_{\text{V}}(\bm{Z})=\frac{1}{n}\sum_{i=1}^{n}\ln\left[\frac{n}{2m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
	\end{equation*}
where $m<n/2$ is a positive integer, $Z_{(i+m)}-Z_{(i-m)}$ is the $m$-spacing and $Z_{(1)}\leq Z_{(2)}\leq\ldots\leq Z_{(n)}$ are the order statistics and $Z_{(i)}= Z_{(1)}$ if $i<1$, $Z_{(i)}= Z_{(n)}$ if $i>n$.

Several authors have explored adaptations to Vasicek's estimator. 
We consider three estimators known for their superior performance&nbsp;\cite{Cassetti2022}:
Correa&nbsp;\cite{correa1995new}, \(\widehat{H}_{\text{C}}\);
Ebrahimi et al&nbsp;\cite{Ebrahimi1994}, \(\widehat{H}_{\mathrm{E}}\);
and Al Omary&nbsp;\cite{IbrahimAlOmari2014}, \(\widehat{H}_{\mathrm{AO}}\).

@@

We use the bootstrap technique to refine the accuracy of existing nonparametric entropy estimators. In this approach, new data sets are generated by replicate sampling from an existing data set&nbsp;\cite{Michelucci2021}.

Let us assume that non-parametric entropy estimators
\(\widehat{H}=\widehat{\theta}(\bm{Z})\) are inherently biased, i.e:\begin{equation}
\label{Eq:bias1}
\operatorname{Bias}\big(\widehat{\theta}(\bm{Z})\big) = E\big[\widehat{\theta}(\bm{Z})\big] - \theta \neq 0.
\end{equation} Our bootstrap-improved estimators are of the form:
\begin{align*}
\widetilde{H} &= 2\widehat{\theta}(\bm{Z}) - \frac{1}{B}\sum_{b=1}^B \widehat{\theta}_b(\bm{Z}^{(b)}),
\end{align*} 
where \(B\) is the number of replications in the bootstrap
technique. 
Applying this methodology, the original estimators of Correa,
Ebrahimi and Al-Omari are now referred to as the proposed bootstrap-enhanced versions:
\(\widetilde{H}_{\text{C}}\),
\(\widetilde{H}_{\text{E}}\), and \(\widetilde{H}_{\text{AO}}\),
respectively.

We analyzed the performance of these estimators with a Monte Carlo study: \(500\) samples of \(G_I^0\) of size
\(n\in\left\{9, 25, 49, 81, 121\right\}\) and different values for mean, number of looks and roughness. We present the results with
\(\mu\in\left\{1, 10\right\}\), \(\alpha=-10\) and \(L=5\). The results are also adherent for the other situations. In the bootstrap technique, each sample is replicated a hundred times with replacement. We use the following heuristic formula for spacing,
\(m=\left[\sqrt{n}+0.5\right]\).

In Figure&nbsp;\ref{fig:Plot_bias_mse_gi0} we show comparisons of bias and mean squared error (MSE) between the original non-parametric entropy estimators and their respective bootstrap-enhanced versions. The use of the bootstrap technique shows higher precision, lower bias and MSE, and improved convergence. The results of the simulation can be found in
Table&nbsp;\ref{tab:table2}.

The precision of the estimators resulting from the comparison of bias and MSE benefits significantly from the bootstrap technique, especially for sample sizes below 81.
